---
title: "PA1_template"
author: "Varun Raste"
date: "23/07/2020"
output: html_document

---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Activity Data Analysis
```{r,message=FALSE,warnings=FALSE,echo=TRUE}
library(readr)
library(dplyr)
df <- read_csv("activity.csv")
attach(df)
df1=df %>% group_by(date)%>% summarise(Total_no_of_steps=sum(steps),Average_no_of_steps=mean(steps),Median_no_of_steps=median(steps))
head(df1)
```

### What is mean total number of steps taken per day?

```{r,message=FALSE,warnings=FALSE,echo=TRUE}
###Total no of steps
df1[,c('date','Total_no_of_steps')]
##Histogram
hist(df1$Total_no_of_steps)
##
df1[,c('date','Average_no_of_steps')]
##
df1[,c('date','Median_no_of_steps')]
```


### What is the average daily activity pattern?
```{r,message=FALSE,warnings=FALSE,echo=TRUE}
df2=df %>% group_by(interval)%>% summarise(Avergae_steps_taken=mean(steps,na.rm=TRUE))
ts.plot(df2$Avergae_steps_taken,xlab="Interval(5 minute)",ylab="Average Steps Taken")
```
 
 100 to 105-minute(1st starts at 12:00 am-12:05 a.m) interval, on average across all the days in the dataset, contains the maximum number of steps.

### For missing values imputation
```{r,message=FALSE,warnings=FALSE,echo=TRUE}
library(questionr)
freq.na(df)
### 
df3=merge(df,df1,by="date",all.x=TRUE)
for(i in 1:length(df3$steps))
{
  df3$Average_no_of_steps[i]=ifelse(is.na(df3$Average_no_of_steps[i])==TRUE,0,df3$Average_no_of_steps[i])
  df3$steps[i]=ifelse(is.na(df3$steps[i])==TRUE,df3$Average_no_of_steps[i],df3$steps[i])
}
head(df3,5)
######
freq.na(df3)

###### Plotting histogram
df$steps=replace(df$steps,is.na(df$steps)==TRUE,0)
df4=df %>% group_by(date)%>% summarise(Total_no_of_steps=sum(steps),Average_no_of_steps=mean(steps),Median_no_of_steps=median(steps))
par(mfrow=c(2,1))
hist(df1$Total_no_of_steps,xlab='Total No. of steps',ylab='Frequency',main='Histogram of Total No. of steps(Before Imputation)')
hist(df4$Total_no_of_steps,xlab='Total No. of steps',ylab='Frequency',main='Histogram of Total No. of steps(After Imputation)')

```


After plotting histogram we can see the distribution near bin 0 has rose. Because we have imputed the NA's. 


### Are there differences in activity patterns between weekdays and weekends?
```{r,message=FALSE,warnings=FALSE,echo=TRUE}
library(ggplot2)
df$Weekday=weekdays(df$date)
df$weekend=ifelse(df$Weekday=='Saturday'|df$Weekday=='Sunday',1,0)
df5=df %>% group_by(weekend,interval) %>% summarise(Average_steps=mean(steps))
ggplot(data = df5, aes(interval, Average_steps)) +
  geom_line(color = "steelblue", size = 1) +
  geom_point(color = "steelblue") + 
  labs(title = "Panel plot for Average no of steps",
       subtitle="1st graph represents for weekdays and later represents weekend",
       y = "Frequency", x = "Interval(5 minute)") + 
  facet_grid(weekend~.) 
```

Yes there is difference for some intervals in weekdays & weekend.Its evident from graph.